Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2015-01-27 Paper/ms.tex
mention posterior prob of existence as calculated from samples

The posterior distribution for the periods $\{P_i\}_{i=1}^N$ is shown in
Figure~\ref{fig:fake_data_periods}. Because of the label switching degeneracy,
the posterior distribution for each period is identical, so we pooled the
samples for all periods. Defining the log-periods by
$S_i =  \log_{10}\left[P_i/(\textnormal{1 day})\right]$,
Figure~\ref{fig:fake_data_periods} is a Monte Carlo representation
of the mixture distribution
\begin{eqnarray}
f(S) &=& \sum_{N=0}^{10} p(N | D)\sum_{i=1}^N p(S_i | N, D).
\end{eqnarray}
If a certain period is accurately measured (i.e. its probability of existence
is close to 1 and the posterior distribution for its period is very narrow)
then it will appear in Figure 3 with a height of $\sim 520$. If the uncertainty
in the period is larger than the histogram bin width then the peak will be
spread over several bins.

@@

It is now recognised that
there are many possible sources for oscillations in a data set and not all
such oscillations should be claimed as planets. 
Our model cannot distinguish between oscillations due to planets and
oscillations due to stellar activity: any oscillations found in the dataset
will be described as ``planets'' by the model. Another important consideration
is the physical stability of the orbital system which is ignored by this type
of analysis \citep{orbital}. However it is interesting that
we find many more signals in the data than previous authors.
By inspecting the posterior distribution for the periods
(Figure~\ref{fig:gliese581_periods}), we see that only four of the periods
are well determined, corresponding to the known periods of Gliese 581 b, c, d,
and e. The other ``periods'' are more uncertain. As with $\nu$ Oph, we can
calculate the posterior probability of any hypothesis
about Gliese 581 by computing the fraction of the posterior samples that have
that property. The posterior probabilities for planets b, c, d, and e, are
close to 1. The posterior probability a signal exists with $\log_{10}(P)
\in [1.55, 1.57]$ is 88\%, and the probability for a signal with
$\log_{10}(P) \in [2.6, 2.8]$ is 85\%.

One possible explanation for the large number of inferred signals that a
non-sinusoidal signal due to stellar activity is being modelled as several
periods \citep[e.g. as happens in asteroseismology when sinusoidal models are
used;][]{astero}, and if the model were extended to include a ``stochastic''
oscillation \citep[e.g.][]{gaussproc}, the number of periods detected may be reduced
substantially. Another contributing factor is the prior for the amplitudes.
With these kinds of models, the posterior distribution for $N$
can also be influenced by the prior for the amplitudes $A_i$. Many authors
assign independent broad priors to the amplitudes, and this causes the
``Occam's razor'' penalty for adding extra signals to be quite strong. Since we
use a hierarchical prior for the amplitudes, if some amplitudes are found to
be low, $\mu_A$ will become small. When $\mu_A$ is small, it is likely that any
extra signals will have small amplitudes, so the ``Occam's razor'' effect is
weaker. An example model fit for Gliese 581 is shown in
Figure~\ref{fig:gliese581}.

>>>>

The posterior distribution for the periods $\{P_i\}_{i=1}^N$ is shown in
Figure~\ref{fig:fake_data_periods}. Because of the label switching degeneracy,
the posterior distribution for each period is identical, so we pooled the
samples for all periods. Defining the log-periods by
$S_i =  \log_{10}\left[P_i/(\textnormal{1 day})\right]$,
Figure~\ref{fig:fake_data_periods} is a Monte Carlo representation
of the mixture distribution
\begin{eqnarray}
f(S) &=& \sum_{N=0}^{10} p(N | D)\sum_{i=1}^N p(S_i | N, D).
\end{eqnarray}
If a certain period is accurately measured (i.e. it appears in close to 100 of
the posterior samples and the distribution for its period is very narrow)
then it will appear in Figure 3 with a height of $\sim 520$. If the uncertainty
in the period is larger than the histogram bin width then the peak will be
spread over several bins.

@@

It is now recognised that
there are many possible sources for oscillations in a data set and not all
such oscillations should be claimed as planets. 
Our model cannot distinguish between oscillations due to planets and
oscillations due to stellar activity: any oscillations found in the dataset
will be described as ``planets'' by the model. Another important consideration
is the physical stability of the orbital system which is ignored by this type
of analysis \citep{orbital}. However it is interesting that
we find many more signals in the data than previous authors.
By inspecting the posterior distribution for the periods
(Figure~\ref{fig:gliese581_periods}), we see that only 4 of the periods
are well determined and have a posterior probability close to 1 (i.e. they
are present in all samples), corresponding to the known periods of Gliese 581
b, c, d, and e. The other ``periods'' are more uncertain.
As with $\nu$ Oph, we can
calculate the posterior probability of any hypothesis
about Gliese 581 by computing the fraction of the posterior samples that have
that property. The posterior probabilities for planets b, c, d, and e, are
close to 1. The posterior probability a signal exists with $\log_{10}(P)
\in [1.55, 1.57]$ is 88\%, and the probability for a signal with
$\log_{10}(P) \in [2.6, 2.8]$ is 85\%.

One possible explanation for the large number of inferred signals that a
non-sinusoidal signal due to stellar activity is being modelled as several
periods \citep[e.g. as happens in asteroseismology when sinusoidal models are
used;][]{astero}, and if the model were extended to include a ``stochastic''
oscillation \citep[e.g.][]{gaussproc}, the number of periods detected may be reduced
substantially. Another contributing factor is the prior for the amplitudes.
With these kinds of models, the posterior distribution for $N$
can be influenced by the prior for the amplitudes $A_i$. Many authors
assign independent broad priors to the amplitudes, and this causes the
``Occam's razor'' penalty for adding extra signals to be quite strong. Since we
use a hierarchical prior for the amplitudes, if some amplitudes are found to
be low, $\mu_A$ will become small. When $\mu_A$ is small, it is likely that any
extra signals will have small amplitudes, so the ``Occam's razor'' effect is
weaker. An example model fit for Gliese 581 is shown in
Figure~\ref{fig:gliese581}.
